---
title: "RWorksheet_Jacildo#4b"
author: "Czharina Mae Jacildo"
date: "2024-10-30"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Using the for loop, create an R script that will display a 5x5 matrix as shown in Figure 1. It must contain vectorA = [1,2,3,4,5] and a 5 x 5 zero matrix.

```{r}
vectorA <- c(1, 2, 3, 4, 5)
matrixx <- matrix (0, nrow = 5, ncol = 5)

for (i in 1:5) {
  for (j in 1:5) {
    matrixx[i, j] <- abs(vectorA[i] - vectorA[j])
  }
}
print(matrixx)
```

2. Print the string "*" using for() function. The output should be the same as shown in Figure

```{r}
rows <- 5

for (i in 1:rows) {
  cat(rep("*", i), "\n")
}
```

3. Get an input from the user to print the Fibonacci sequence starting from the 1st input up to 500. Use repeat and break statements. Write the R Scripts and its output.

```{r}
a <- 0
b <- 1

c <- readline(prompt = "Enter a number: ")

repeat 
  c <- a + b
  if (c > 500) break
  if (c >= start) {
  a <- b
  b <- c
  print (c)
}
```

4. Import the dataset as shown in Figure 1 you have created previously.
a. What is the R script for importing an excel or a csv file? Display the first 6 rows of the dataset? Show your codes and its result
```{r}

```

