---
title: "RWorksheet_Jacildo#4a"
author: "Czharina Mae Jacildo"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.

``` {r}
shoe_size <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)

height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)

gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")

data <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)
print(data)
```

a. Describe the data.
- It shows the data about shoe size, height, and gender.

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.

``` {r}
male <- subset(data, gender == "M")
female <- subset(data, gender == "F")

print(male)
print(female)
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
``` {r}
shoe_size_mean <- mean(shoe_size)
height_mean <- mean(height)

print(shoe_size_mean)
print(height_mean)
```

d. Is there a relationship between shoe size and height? Why?
- Yes, usually taller people have longer feet.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.

``` {r}
months_vector <- c("March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August", "July","December","August","August","September","November","February","April")

factor_months_vector <- factor(months_vector)
print(factor_months_vector)
```

3. Then check the summary() of the months_vector and factor_months_vector. | Inter-pret the results of both vectors. Are they both equally useful in this case?

``` {r}
summary_months_vector <- summary(months_vector)
summary_factor_months <- summary(factor_months_vector)

print(summary_months_vector)
print(summary_factor_months)
```

4. Create a vector and factor for the table below.
``` {r}
direction <- c("East", "West", "North")
frequency <- c(1, 4, 3)

new_order_data <- factor(direction,levels = c("East","West","North"))
print(new_order_data)
```

5. Enter the data below in Excel with file name = import_march.csv

a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

``` {r}
library(readxl)
excelData <- read.table("/cloud/project/worksheet#4/import_march.csv", header = TRUE, sep = ",")
```

b. View the dataset. Write the R scripts and its result.
``` {r}
print(excelData)
```