---
title: "RWorksheet_Jacildo#4a"
author: "Czharina Mae Jacildo"
date: "2024-10-18"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. The table below shows the data about shoe size and height. Create a data frame.

``` {r}
shoe_size <- c(6.5, 9.0, 8.5, 8.5, 10.5, 7.0, 9.5, 9.0, 13.0, 7.5, 10.5, 8.5, 12.0, 10.5, 13.0, 11.5, 8.5, 5.0, 10.0, 6.5, 7.5, 8.5, 10.5, 8.5, 10.5, 11.0, 9.0, 13.0)

height <- c(66.0, 68.0, 64.5, 65.0, 70.0, 64.0, 70.0, 71.0, 72.0, 64.0, 74.5, 67.0, 71.0, 71.0, 77.0, 72.0, 59.0, 62.0, 72.0, 66.0, 64.0, 67.0, 73.0, 69.0, 72.0, 70.0, 69.0, 70.0)

gender <- c("F", "F", "F", "F", "M", "F", "F", "F", "M", "F", "M", "F", "M", "M", "M", "M", "F", "F", "M", "F", "F", "M", "M", "F", "M", "M", "M", "M")

data <- data.frame(Shoe_size = shoe_size, Height = height, Gender = gender)
print(data)
```

a. Describe the data.
- It shows the data about shoe size, height, and gender.

b. Create a subset by males and females with their corresponding shoe size and height.
What its result? Show the R scripts.

``` {r}
male <- subset(data, gender == "M")
female <- subset(data, gender == "F")

print(male)
print(female)
```

c. Find the mean of shoe size and height of the respondents. Write the R scripts and its
result.
``` {r}
shoe_size_mean <- mean(shoe_size)
height_mean <- mean(height)

print(shoe_size_mean)
print(height_mean)
```

d. Is there a relationship between shoe size and height? Why?
- Yes, usually taller people have longer feet.

2. Construct character vector months to a factor with factor() and assign the result to factor_months_vector. Print out factor_months_vector and assert that R prints out
the factor levels below the actual values.

``` {r}
months_vector <- c("March","April","January","November","January",
"September","October","September","November","August",
"January","November","November","February","May","August", "July","December","August","August","September","November","February","April")

factor_months_vector <- factor(months_vector)
print(factor_months_vector)
```

3. Then check the summary() of the months_vector and factor_months_vector. | Inter-pret the results of both vectors. Are they both equally useful in this case?

``` {r}
summary_months_vector <- summary(months_vector)
summary_factor_months <- summary(factor_months_vector)

print(summary_months_vector)
print(summary_factor_months)
```

4. Create a vector and factor for the table below.
``` {r}
direction <- c("East", "West", "North")
frequency <- c(1, 4, 3)

new_order_data <- factor(direction,levels = c("East","West","North"))
print(new_order_data)
```

5. Enter the data below in Excel with file name = import_march.csv

a. Import the excel file into the Environment Pane using read.table() function.
Write the code.

``` {r}
library(readxl)
excelData <- read.table("/cloud/project/worksheet#4/import_march.csv", header = TRUE, sep = ",")
```

b. View the dataset. Write the R scripts and its result.
``` {r}
print(excelData)
```

6. Full Search
a. Create an R Program that allows the User to randomly select numbers from 1 to 50. Then display the chosen number. If the number is beyond the range of the selected choice, it will have to display a string “The number selected is beyond the range of 1 to 50”. If number 20 is inputted by the User, it will have to display “TRUE”, otherwise display the input number.

```{r}
user <- readline(prompt = "Enter a number from 1 to 50: ")

if (user < 1 || user > 50) {
  cat("The number selected is beyond the range of 1 to 50\n")
} else if (user == 20) {
  cat("TRUE\n")
} else {
  print(user)
}
```

7. Change
a. Write a function that prints the minimum number of bills that must be paid, given the
price of the snack.
Input: Price of snack (a random number divisible by 50) Output: Minimum number of bills needed to purchase a snack.

```{r}
min_bills <- function(price) {
  bills <- c(1000, 500, 200, 100, 50)
  
  count <- 0
  
  # Loop through each bill denomination
  for (bill in bills) {
    # Calculate how many bills of this denomination can be used
    while (price >= bill) {
      price <- price - bill
      count <- count + 1
    }
  }
  
  # Print the result
  cat("Minimum number of bills needed:", count, "\n")
}

# Example usage
min_bills(250)
```

8. The following is each student’s math score for one semester. Based on this, answer the
following questions.

a. Create a dataframe from the above table. Write the R codes and its output.

```{r}
name <- c("Annie", "Thea", "Steve", "Hanna")
grade1 <- c(85, 65, 75, 95) 
grade2 <- c(65, 75, 55, 75)
grade3 <- c(85, 90, 80, 100)
grade4 <- c(100, 90, 85, 90)

sem <- data.frame(Name = name, Grade1 = grade1, Grade2 = grade2, Grade3 = grade3, Grade4 = grade4)

print(sem)
```

b. Without using the rowMean function, output the average score of students whose average math score over 90 points during the semester. write R code and its output.
Example Output: Annie’s average grade this semester is 88.75.

```{r}

```

c. Without using the mean function, output as follows for the tests in which the average score was less than 80 out of 4 tests. Example output: The nth test was difficult.
```{r}

```

d. Without using the max function, output as follows for students whose highest score for a semester exceeds 90 points.
Example Output: Annie’s highest grade this semester is 95.
```{r}

```

