---
title: "RWorksheet_Jacildo#2"
author: "Czharina Mae Jacildo"
date: "2024-09-25"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. Create a vector using : operator

a. Sequence from -5 to 5. Write the R code and its output.
```{r}
a <- c(-5:5)
print(a)
```
Describe its output.

- The output shows a sequence from -5 to 5.

b. What will be the value of x?
```{r}
x <- 1:7
print(x)
```
- numbers from one to seven.

2. Create a vector using seq() function

a. seq(1, 3, by=0.2) # specify step size
Write the R script and its output. 
```{r}
vecseq <- seq(1, 3, by=0.2)
print(vecseq)
```
Describe the output.

- The sequence increases by 0.2 from number one until three.

3. 
```{r}
ages <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29, 35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 43, 53, 41, 51, 35, 24, 33, 41, 53, 40, 18, 44, 38, 41, 48, 27, 39, 19, 30, 61, 54, 58, 26, 18)
```
a. Access 3rd element, what is the value?
```{r}
ages[3]
```

b. Access 2nd and 4th element, what are the values?
```{r}
ages[c(2, 4)]
```

c. Access all but the 4th and 12th element is not
included. Write the R script and its output.
```{r}
ages[-c(4, 12)]
```

4. Create a vector x <- c("first"=3, "second"=0, "third"=9). Then named the
vector, names(x).

a. Print the results. Then access x[c("first", "third")].

b. Write the code and its output.
```{r}
x <- c("first"=3, "second"=0, "third"=9)
names(x)
x[c("first", "third")]
```
Describe the output

- The first and third in the vector and its value is displayed with spaces.

5. Create a sequence x from -3:2.

a. Modify 2nd element and change it to 0;
Describe the output.

- The second element was changed to 0.

b. Write the code and its output.
```{r}
x <- -3:2
x[2] <- 0
x
```

6. The following data shows the diesel fuel purchased by Mr. Cruz.

a. Create a data frame for month, price per liter (php)
and purchase-quantity (liter). Write the R scripts and
its output.
```{r}
month = c("Jan", "Feb", "Mar", "Apr", "May", "June")
price_per_liter = c(52.50, 57.25, 60.00, 65.00, 74.25, 54.00)
purchase_quantity = c(25, 30, 40, 50, 10, 45)

df <- data.frame(Month = month, PricePerLiter = price_per_liter, PurchaseQuantity = purchase_quantity)
print(df)
```

b. What is the average fuel expenditure of Mr. Cruz from
Jan to June? Note: Use ‘weighted.mean(liter,
purchase)‘. Write the R scripts and its output.
```{r}
ave_fe <- weighted.mean(df$Price_per_liter, df$Purchase_quantity)
print(ave_fe)
```

7. R has actually lots of built-in datasets. For example, the rivers data “gives the lengths
(in miles) of 141 “major” rivers in North America, as compiled by the US Geological
Survey”.

a. Type “rivers” in your R console.
Create a vector data with 7
elements

b. What are the results?
- The length, sum, mean, median, variance, standard deviation, minimum, maximum of rivers.

c. Write the R scripts and its outputs.
```{r}
data <- c(length(rivers), sum(rivers), mean(rivers), median(rivers), var(rivers), sd(rivers), min(rivers), max(rivers))
print(data)
```

8. The table below gives the 25 most powerful celebrities and their annual pay as ranked
by the editions of Forbes magazine and as listed on the Forbes.com website.

a. Create vectors according to the above table.
Write the R scripts and its output.
```{r}
powerRanking <- c(1, 2, 3, 4, 5, 6, 7, 8, 9, 10, 11, 12, 13, 14, 15, 16, 17, 18, 19, 20, 21, 22, 23, 24, 25)

celebrityName <- c("Tom Cruise", "Rolling Stones", "Oprah Winfrey", "U2", "Tiger Woods", "Steven Spielberg", "Howard Stern", "50 Cent", "Cast of the Sopranos", "Dan Brown", "Bruce Springsteen", "Donald Trump", "Muhammad Ali", "Paul McCartney", "George Lucas", "Elton John", "David Letterman", "Phil Mickelson", "J.K. Rowling", "Bradd Pitt", "Peter Jackson", "Dr. Phil McGraw", "Jay Leno", "Celine Dion", "Kobe Bryant")

pay <- c(67, 90, 225, 110, 90, 332, 302, 41, 52, 88, 55, 44, 55, 40, 233, 34, 40, 47, 75, 25, 39, 45, 32, 40, 31)

celebrities <- data.frame(PowerRanking = powerRanking, Celebrity = celebrityName, Pay = pay)
print(celebrities)
```
b. Modify the power ranking and pay of J.K. Rowling.
Change power ranking to 15 and pay to 90. Write the
R scripts and its output.
```{r}
powerRanking[19] <- 15
pay[19] <- 90
celebrities <- data.frame(PowerRanking = powerRanking, Celebrity = celebrityName, Pay = pay)
print(celebrities)
```

c. Create an excel file from the table above and save it
as csv file(PowerRanking). Import the csv file into
the RStudio. What is the R script?
```{r}
write.csv(celebrities, file = "PowerRanking.csv", row.names = FALSE)
importedData <- read.csv("PowerRanking.csv")
print(importedData)
```
d. Access the rows 10 to 20 and save it as Ranks.RData.
Write the R script and its output.
```{r}
ranksData <- celebrities[10:20, ]
save(ranksData, file = "Ranks.RData")
print(ranksData)
```

e. Describe its output.

In a.) Data frame containing the celebrity names, their power rankings, and their pay.

In b.) After modifying J.K. Rowling’s power ranking and pay, the new values of power ranking is 15 and pay is 90.

In c.) The csv file, will be saved with the table's data. Importing it will result in a similar data frame to the original one.

In d.) shows celebrities ranked from 10th to 20th along with their pay.

9. Download the Hotels-Vienna https://tinyurl.com/Hotels-Vienna
a. Import the excel file into your RStudio.
What is the R script?
```{r}
library(readxl)
hotelsData <- read_excel("hotels-vienna.xlsx")
print(head(hotelsData))
```

b. How many dimensions does the dataset have?
What is the R script? What is its output?
```{r}
dimensions <- dim(hotelsData)
print(dimensions)
```

c. Select columns country, neighbourhood,
price, stars, accommodation_type, and
ratings. Write the R script.
```{r}
class(hotelsData)
selection <- hotelsData[,c("country", "neighbourhood", "price", "stars", "accommodation_type", "rating")]
print(selection)
```

d. Save the data as **new.RData to your RStudio. Write the
R script.
```{r}
save(selection, file = "new.RData")
print(head(selection))
print(tail(selection))
```

e. Display the first six rows and last six rows of the
new.RData. What is the R script?
```{r}
load("new.RData")
print(head(selection))
print(tail(selection))
```

10. Create a list of ten (10) vegetables you ate during your lifetime. If none, just list down.

a. Write the R scripts and its output.
```{r}
vegetables <- c("Squash", "BitterGourd", "Potato", "Pechay", "SweetPotato", "Corn", "BeanSprouts", "Sayote", "Mongo", "Upo")
print(vegetables)
```

b. Add 2 additional vegetables after the last vegetables
in the list. What is the R script and its output?
```{r}
vegetables <- c(vegetables, "radish", "KangKong")
print(vegetables)
```
c. Add 4 additional vegetables after index 5. How many
datapoints does your vegetable list have? What is the
R script and its output?
```{r}
vegetables <- append(vegetables, c("BaguioBeans", "Malunggay", "Garlic", "Onion"))
print(vegetables)
vegdatapoints <- length(vegetables)
print(vegdatapoints)
```
d. Remove the vegetables in index 5, 10, and 15. How
many vegetables were left? Write the codes and its
output.
```{r}
vegetables <- vegetables[-c(5, 10, 15)]
print(vegetables)
left <- length(vegetables)
print(left)
```