---
title: "RWorksheet_Jacildo#1"
author: "Czharina Mae Jacildo"
date: "2024-09-04"
output: pdf_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

1. 
a. How many data points?
- There are 34 data points.

b. Write the R code and its output.
```{r}
age <- c(34, 28, 22, 36, 27, 18, 52, 39, 42, 29,
35, 31, 27, 22, 37, 34, 19, 20, 57, 49, 50, 37, 46, 25, 17, 37, 42, 53, 41, 51, 35, 24, 33, 41.)
points <- length(age)
print(points)
```

2. Find the reciprocal of the values for age.
Write the R code and its output.
```{r}
reciprocal_age <- 1 / age
print(reciprocal_age)
```
3. What happen to the new_age?
- It creates a new vector combining 2 age vectors with 0 in the middle.
```{r}
new_age <- c(age, 0, age)
print(new_age)
```
4. Sort the values for age.
```{r}
age <- sort(age)
print(age)
```
5. Find the minimum and maximum value for age.
```{r}
minimum_age <- min(age)
maximum_age <- max(age)
print(minimum_age)
print(maximum_age)
```
6. How many data points?
- There are 12 data points.

b. Write the R code and its output.
```{r}
data <- c(2.4, 2.8, 2.1, 2.5, 2.4, 2.2, 2.5, 2.3, 2.5, 2.3, 2.4, 2.7)
points <- length(data)
print(points)

```
7. Generates a new vector for data where you double every value of the data. | What
happen to the data?
- The elements in the data vector were multiplied by two.
```{r}
double_nv <- data * 2
print(double_nv)

```
8.1 Integers from 1 to 100.
```{r}
integers <- seq(1, 100)
print(integers)
```
8.2 Numbers from 20 to 60
```{r}
numbers <- seq(20, 60)
print(numbers)
```
*8.3 Mean of numbers from 20 to 60
```{r}
mean_numbers <- mean(20:60)
print(mean_numbers)
```
*8.4 Sum of numbers from 51 to 91
```{r}
sum_numbers <- sum(51:91)
print(sum_numbers)
```
*8.5 Integers from 1 to 1,000
```{r}
ints <- seq(1:1000)
print(ints)
```
a. How many data points from 8.1 to 8.4?
- 143 data points

b. Write the R code and its output from 8.1 to 8.4.
```{r}
dp1 <- length(integers) 
dp2 <- length(numbers) 
dp3 <- length(mean_numbers) 
dp4 <- length(sum_numbers) 
data_points <- sum(dp1+dp2+dp3+dp4)
print(data_points)
```
c. For 8.5 find only maximum data points until 10.
```{r}
m <- 1:1000
mx <- max(m[t <- 10])
print(mx)
```
9. Print a vector with the integers between 1 and 100 that are not divisible by 3, 5 and 7 using filter option.
```{r}
Filter(function(i) { all(i %% c(3,5,7) != 0) }, seq(100))
```
10. Generate a sequence backwards of the integers from 1 to 100.
Write the R code and its output.
```{r}
backward <- rev(1:100)
print(backward)
```
11. List all the natural numbers below 25 that are multiples of 3 or 5.
Find the sum of these multiples.
```{r}
nums <- 1:24
multiples <- nums[nums%%3==0|nums%%5==0]
print(nums)

sum_multiples <- sum(multiples)
print(sum_multiples)
```

a. How many data points from 10 to 11?
- 12 data points.

b. Write the R code and its output from 10 and 11.
```{r}
dp10 <- length(multiples)
dp11 <- 1
total <- dp10 + dp11
print(total)
```
12. Statements can be grouped together using braces ‘{’ and ‘}’. A group of statements
is sometimes called a block. Single statements are evaluated when a new line is typed
at the end of the syntactically complete statement. Blocks are not evaluated until a
new line is entered after the closing brace.
```{r}
x <- {0 + x + 5 +}
```
Describe the output.
- Error, it is incomplete

13. *Set up a vector named score, consisting of 72, 86, 92, 63, 88, 89, 91, 92, 75, 75 and 77. To access individual elements of an atomic vector, one generally uses the x[i] construction.
Find x[2] and x[3]. Write the R code and its output.
```{r}
score <- c(72, 86, 92, 63, 88, 89, 91, 92, 75,
75, 77)
x2 <- score[2]
x3 <- score[3]
print(x2)
print(x3)
```

14. *Create a vector a = c(1,2,NA,4,NA,6,7).
```{r}
a <- c(1,2,NA,4,NA,6,7)
print(a)
```
a. Change the NA to 999 using the codes print(a,na.print="-999").
```{r}
print(a,na.print="-999")
```
b. Write the R code and its output. Describe the output.

15. A special type of function calls can appear on the left hand side of the assignment
operator as in > class(x) <- "foo".
```{r}
name = readline(prompt="Input your name: ")
age = readline(prompt="Input your age: ")
print(paste("My name is",name, "and I am",age ,"years old."))
print(R.version.string)
```
What is the output of the above code?
"My name is (name) and I am (age) years old." 
